Novelty Detection Using Graphical Models for Semantic Room Classification

### 2011-05-16 semantic-nd/article.tex
Changes in introduction.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
There has been several efforts in the area of Artificial Intelligence and Robotics in creating
robots that are able to interact with humans and their environments.
One of the existing problems is a reliable high-level localization method that can be deployed
into new and unknown environments.
This article focus on the mapping, using \emph{semantic data}, of indoor environments such as
houses and offices to room categories such as kitchen, corridor, office.
And how to perform novelty detection on them: \emph{detect a new room category that was not
present in the labelled data}.

Dora\cite{dora} (CogX: Dora) was used as a base system where to implement the presented novelty
detection system.
As Dora moves through the environment its \emph{conceptual layer} builds a structural and
probabilistic representation of the space instantiated as a \emph{graphical model}.
That model connects the sensed properties together with the variables used to model the world.
And allows to perform queries on the probabilities of aspects of the environment.
For example: where is most likely to find a cereal box\cite{exploiting}.

The ability to recognize novel categories on the modelled variables would allow to increase
reliability and allow the creation of self-extending behaviours.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
There has been several efforts in the area of Artificial Intelligence and Robotics
in creating robots that are able to interact with humans and their environments.
One of the existing problems is a reliable high-level localization method that can
be deployed into new and unknown environments or where its impossible to access a
previously labelled map of the area. Nonetheless the need of reliability and ability
to deal with new cases also arises due to dynamic changes on the environment through
time and with human action.
This article focus on the mapping, using \emph{semantic data}, of indoor
environments such as houses and offices to room categories such as kitchen,
corridor, office.

Due to unknown and dynamic nature of the operating environment of the agent the
mapping methods should be resilient to new room categories it has never seen
before. This is: perform novelty detection on the concepts the agent is aware of.

Dora\cite{dora} (CogX: Dora) was used as a base system where to implement the
presented novelty detection system as it already possesses a developed architecture
oriented on dealing with probabilistic and semantic information and performing
reasoning over it.
Implementing novelty detection capabilities on an agent such as Dora, would allow to
to increase reliability and allow the creation of self-extending behaviours.

Its with a final interest on integrated the developed methods on Dora that this
initial work focus on the novelty detection using probabilistic graphical models
that model the conceptual knowledge and information handled by Dora.

@@

As Dora moves through the environment its \emph{conceptual layer} builds a structural and
probabilistic representation of the space instantiated as a \emph{graphical model}.
That model connects the sensed properties together with the variables used to model the world.
And allows to perform queries on the probabilities of aspects of the environment.
For example: where is most likely to find a cereal box\cite{exploiting}.
